### 2016-02-17 tex/paper2/paper2.tex

%%
%% Beginning of file 'sample.tex'
%%
%% Modified 2005 December 5
%%
%% This is a sample manuscript marked up using the
%% AASTeX v5.x LaTeX 2e macros.

%% The first piece of markup in an AASTeX v5.x document
%% is the \documentclass command. LaTeX will ignore
%% any data that comes before this command.

%% The command below calls the preprint style
%% which will produce a one-column, single-spaced document.
%% Examples of commands for other substyles follow. Use
%% whichever is most appropriate for your purposes.
%%
%%\documentclass[12pt,preprint]{aastex}

%% manuscript produces a one-column, double-spaced document:

\documentclass[onecolumn]{emulateapj}
%\documentclass[twocolumn]{aastex}
%\documentclass[referee]{aastex}
\usepackage{graphicx}
\usepackage{color}
\usepackage{multirow}
\usepackage{setspace}

%\usepackage{showkeys}
\usepackage{epsfig}
\usepackage{natbib}
\usepackage{wasysym}
%\usepackage{subfigure}
\usepackage{verbatim}

\bibliographystyle{apj}	%% AASTeX

%% preprint2 produces a double-column, single-spaced document:

%% \documentclass[preprint]{aastex}

%% Sometimes a paper's abstract is too long to fit on the
%% title page in preprint2 mode. When that is the case,
%% use the longabstract style option.

%% \documentclass[preprint2,longabstract]{aastex}

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
%% If you are submitting to a journal that translates manuscripts
%% into SGML, you need to follow certain guidelines when preparing
%% your macros. See the AASTeX v5.x Author Guide
%% for information.

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{jack.ireland@nasa.gov}
\newcommand{\PS}{power spectrum}
\newcommand{\mPS}{mean power spectrum}

\newcommand{\PA}{power spectra}
\newcommand{\mPA}{mean power spectra}

\newcommand{\PL}{power-law}
\newcommand{\PLs}{power-laws}

\newcommand{\Fps}{Fourier \PS}
\newcommand{\mFps}{mean Fourier \PS}

\newcommand{\Fpa}{Fourier \PA}
\newcommand{\mFpa}{mean Fourier \PA}

\newcommand{\allaiaeuv}{171\AA, 193\AA, 211\AA, 335\AA, 94\AA\ and 131\AA}

\newcommand{\BF}{ }

%% You can insert a short comment on the title page using the command below.

%\slugcomment{Not to appear in Nonlearned J., 45.}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.
%% Running heads will not print in the manuscript style.

\shorttitle{Spatial variation of coronal Fourier power spectra}
\shortauthors{Ireland et al.}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Spatial variation of coronal Fourier power spectra}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{J. Ireland}
\affil{ADNET Systems, Inc.,NASA Goddard Space Flight Center, MC. 671.1, Greenbelt, MD 20771, USA.}

\author{R. T. J. McAteer}
\affil{Department of Astronomy, New Mexico State University, Las
  Cruces, NM.}

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.

\begin{abstract}
The dynamics of the solar corona are investigated using data from the
six extreme ultraviolet channels (EUV) of the  Atmospheric Imaging
Assembly (AIA) of the Solar Dynamics Observatory (SDO).

It is shown that in almost all locations the \Fps\ are well be described by
a \PL\ at lower frequencies that tails to flat spectrum at higher
frequencies, plus a Gaussian-shaped contribution that varies depending
on the region studied.  The spatial distribution of the power law
index is described as a function of AIA EUV channel.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. The uncommented
%% example has been keyed in ApJ style. See the instructions to authors
%% for the journal to which you are submitting your paper to determine
%% what keyword punctuation is appropriate.

\keywords{Sun: corona, Sun: oscillations, methods: data analysis,
  methods: statistical}

%% From the front matter, we move on to the body of the paper.
%% In the first two sections, notice the use of the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below. We have
%% chosen the first three characters of the first author's name plus
%% the last two numeral of the year of publication as our KEY for
%% each reference.

%% Authors who wish to have the most important objects in their paper
%% linked in the electronic edition to a data center may do so by tagging
%% their objects with \objectname{} or \object{}.  Each macro takes the
%% object name as its required argument. The optional, square-bracket 
%% argument should be used in cases where the data center identification
%% differs from what is to be printed in the paper.  The text appearing 
%% in curly braces is what will appear in print in the published paper. 
%% If the object name is recognized by the data centers, it will be linked
%% in the electronic edition to the object data available at the data centers  
%%
%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
%% the brackets must be escaped with backslashes when used in the first
%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
%%  Otherwise, LaTeX will issue an error.

\section{Introduction}\label{sec:int}
Coronal seismology is the study of oscillatory phenomena in the solar
corona.  First suggested by \cite{1970PASJ...22..341U}, the practice
of coronal seismology attempts to locate oscillatory phenomena in the
solar corona, identify the wave modes present, and then use
theoretical descriptions of those wave modes to infer the physical
conditions of the solar corona.  The practical application of coronal
seismological ideas began in earnest with observations of wave modes
{\BF described in \cite{1998ApJ...501L.217D} and
  \cite{1999SoPh..186..207B} from data captured by Extreme Ultraviolet
  Telescope (EIT, \citealp{1995SoPh..162..291D}) on board the Solar and Heliospheric
  Observatory (SOHO, \citealp{1995SoPh..162....1D}), and also by
  \citep{1999ApJ...520..880A} using data from the Transition Region
  and Coronal Explorer (TRACE, \citealp{1999SoPh..187..229H}).  Since these initial works,
  many papers have been published describing oscillatory phenomena in
  the corona and their interpretation in terms of theoretically known
  wave modes.  The review articles \cite{lrsp-2005-3} and
  \cite{2012RSPTA.370.3193D} refer to many more articles on both
  observational and theoretical coronal seismology}.

Oscillatory phenomena in the solar corona are relatively rare, both
spatially and temporally.  Two types of oscillatory motions have
received a lot of attention.  The first is known as a transverse
oscillation, since the loop motion is approximately transverse to the
extent of the loop \citep{1999Sci...285..862N}.  These motions appear
to be triggered by nearby flaring events, and are identified with the
fundamental harmonic of the kink mode of the flux tube that comprises
the coronal loop.  The second type is known as a longitudinal
oscillation, or propagating disturbance (PD).  These waves propagate
along flux tubes, and have been identified as slow mode
magnetohydrodynamic (MHD) magneto-acoustic waves (although alternative
interpretations do exist that do not invoke oscillatory phenomena, for
example, \citealp*{0004-637X-722-2-1013} and
\citealp*{2041-8205-727-2-L37}).

\cite{ireland2015} analyzed power spectra summed over physically
different extended regions.  It was noted that some regions, when
observed in two different AIA channels result in different power-law
power indices, while other regions have approximately similar
power-law spectral indices.  For example, loop footpoint locations had
very similar spectral index of approximately 2.28, whereas AIA 171\AA\
and 193\AA\ had spectral indices of 1.76 and 2.05 respectively.  It is
therefore clear that different regions exhibit different evolutionary
behavior.  This paper extends that work by fitting models of power
spectra arising from time-series in all six EUV channels (\allaiaeuv).
This will result in more detailed infomation on the spatial
distribution of Fourier power in the AIA EUV channels.

Section \ref{sec:obs} describes the observations used, while Section
\ref{sec:anal} describes the method of analysis employed and presents
the results.  Section \ref{sec:discuss} discusses the detected
oscillatory content, and the implications of the observed power-law
power spectra.  Section \ref{sec:conc} summarizes the discussion and
outlines future work.

\section{Observations}\label{sec:obs}
The region studied is the same as that used in the previous
\cite{ireland2015} study.  The selected area covers a single
$\beta\gamma$ active region, NOAA AR 11575.  As was noted in
\cite{ireland2015} , this area is quiescent in that there were no
X-ray flares or other large-scale disturbances (for example, filament
eruptions or rearrangments of long bright coronal loops) over the
spatial extent of the cutout within the time range.
\cite{ireland2015} note the power spectra

{\BF AIA} observes images of the Sun in multiple channels
simultaneously and continuously, at high cadence (12 seconds) and at
high spatial resolution (0.6 arcseconds per pixel).  {\BF AIA} data
was acquired using the Lockheed Martin Solar Astrophysics cutout
service at {\it http://lmsal.com/get\_aia\_data/}.  Six hours of image
data in the time range 2012-09-23 00:00 - 06:00 UT, at the full 12
second cadence, in EUV wavebands \allaiaeuv\ are used.  The data in
each AIA channel are co-aligned to remove solar rotation using the
process as described in Appendix \ref{sec:app:data}.  The result are a
set of datacubes that show the same portion of the Sun for the
duration of the data.

\section{Analysis}\label{sec:anal}
We wish to analyze the frequency content of these resulting datacubes
as a function of spatial location.  Figure \ref{fig:loc171193} shows
the location of the region considered. This region was chosen because
it contains many examples of commonly observed structures in the
corona - moss, quiet Sun, sunpots and loop footpoints.
\begin{figure}
\centerline{
\plottwo{shutdownfun3_6hr_disk_1.5_171.eps}{shutdownfun3_6hr_disk_1.5_193.eps}
}
\caption{Locations of the regions analyzed in each of AIA 171\AA\ and
  193\AA, overlaid on sample images taken halfway through the six hour
  span used to form the time-series analyzed.  The spatial extent of
  the data is (-501'', -138'') in the x-direction, and (-80'', 107'')
  in the y-direction, in heliocentric cartesian co-ordinates.  The
  images show NOAA AR 11575 and some surrounding area.}
\label{fig:loc171193}
\end{figure}

Consider the emission from a single pixel, $I(t)$.  The Fourier power
spectrum for this emission was calculated as follows.  First, the
time-series was normalized by calculating $(I(t) - \langle I(t)
\rangle)/\langle I(t) \rangle$. Secondly, the normalized time-series
was apodized using the Hanning window, reducing ringing effects in the
\PS.  The full 12 second cadence is used, yielding time series of 1800
samples.  The full spatial resolution is also retained in order to
obtain fine detail on the variation of frequency content as a function
of location.  Figure \ref{fig:compare171193} show example power
spectra, plotted using a log-log scale, from a single location inside
each of the regions shown in Figure \ref{fig:loc171193}.  The shape of
the power spectra indicate that the Fourier power approximately
follows a \PL\ at lower frequencies, and flattens out at higher
frequencies.

There is evidence that these \PLs\ extend to much longer
frequencies than those considered here, down to approximately 0.01
mHz. \cite{2014AA...563A...8A} integrate emission over small portions
of active regions and the quiet Sun as observed in the 195\AA\ passband
images from EIT and show that the resulting time-series has an
approximate \PL\ \Fps\ over the frequency range 0.01 - 1 mHz.

\begin{figure}
\centerline{
\plottwo{loopfootpoints.centralpower.eps}{moss.centralpower.eps}
}
\centerline{
\plottwo{qs.centralpower.eps}{sunspot.centralpower.eps}
}
\caption{Log-log plot of the Fourier power of a time-series at the AIA
  pixel overlying a randomly chosen pixel in each of the regions
  considered (see Figure \protect\ref{fig:loc171193}).  The vertical
  black lines indicate the 3 and 5 minute frequencies.}
\label{fig:compare171193}
\end{figure}

\begin{figure}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_loopfootpoints.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_loopfootpoints.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_moss.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_moss.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_qs.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_qs.hanning.relative.power_spectra_distributions.eps}
}
\centerline{
\epsscale{0.8}\plottwo{shutdownfun3_6hr_disk_1.5_171_sunspot.hanning.relative.power_spectra_distributions.eps}{shutdownfun3_6hr_disk_1.5_193_sunspot.hanning.relative.power_spectra_distributions.eps}
}
\caption{{\BF The proportion of the number of pixels in each image
    that have a given Fourier power at some sample Fourier
    frequencies.  The left hand column gives results for AIA 171\AA,
    and the right hand column, AIA 193\AA.  Regions are indicated in
    each plot title.}}
\label{fig:dist171193}
\end{figure}
Figure \ref{fig:dist171193} shows the probability distribution of the
Fourier power of the pixels in each region at five selected
frequencies, for all the pixels in each region.  These distributions
are centrally peaked, show a slightly heavier tail towards lower
Fourier powers (negative skew), are slightly broader when compared to
a Gaussian distribution, and are all unimodal.  These observations
guide the choice of how to average the \PS\ over each of the selected
region.  The nature of these distributions suggest that the arithmetic
mean of the Fourier power is biased by the very largest powers.
Therefore, the mean value of the logarithm of the Fourier power is a
better description of the range of values in the data (this is
equivalent to the geometric mean of the Fourier power).  The \PA\ used
below are calculated from the mean value of the logarithm of the
Fourier power

\subsection{Results}\label{ssec:results}
Each Fourier power spectrum is fit using both models $M_{1}$ and
$M_{2}$.  The $\chi^{2}$-analog is calculated for the best fit to each
observed power spectrum.

A good fit is found if (a) the fitting routine does not report an
error, (b) the $\chi^{2}$-analog lies within the 95\% confidence
interval (see Appendix \ref{sec:app:chi}) and (c) all the parameters
found by the fitting process lie inside acceptable limits (see
Appendix \ref{sec:app:parlimits}).  This qualifies model fits as
worthy of further consideration.  The Bayesian Information Criterion
(BIC) is used to express a preference of one model over another.

The model $M_{k}$ is preferred if
\begin{equation}
BIC_{k} - BIC_{j} \le -6, \forall j\ne k.
\end{equation}

This process selects decides which model and

Figure \ref{fig:fit171193} show the fit results of each model to the
data.  For details on the fitting procedure, see Appendix
\ref{sec:app:ind}.
{\BF The two models are assessed for how well they describe the data in two
different ways. Model selection is implemented using the Akaike Information Criterion
(AIC; \citealp*{akaike}).  The AIC measures the amount of information
lost when fitting a model to data \citep{2000Ap&SS.271..213T}.  The
AIC is defined as
\begin{equation}\label{eqn:aic}
AIC \equiv 2k -2 \ln L_{max}
\end{equation}
where $L_{max}$ is the maximum likelihood achievable by the model, and
$k$ is the number of parameters in the model ($k=3$ for Model 1, and
$k=6$ for Model 2).  Model selection is guided by calculating
$\triangle AIC = AIC_{1} - AIC_{2}$.  Positive values of $\triangle
AIC$ indicate that Model 2 loses less information compared to Model 1,
leading to a preference for Model 2 over Model 1
\citep{2007MNRAS.377L..74L}.  The AIC indicates that Model 2 is
overwhelmingly preferred to Model 1 (see Figure \ref{fig:fit171193}).}
{\BF The AIC reveals which (out of a set of models) is preferred.  It
  does not give an assessment of the goodness-of-fit of each model.
  Goodness-of-fit is estimated by calculating reduced-$\chi^{2}$,
  $\chi^{2}_{r}=\sum_{j=1}^{N}\left[(P_{i}(\nu_{j})-D_{j})/\sigma_{j}^{2}\right]/(N-k-1)$,
  where $N=899$ is the number of frequencies, $P_{i}(\nu_{j})$ is the
  model power $P_{i}$ at frequency $\nu_{j}$, $D_{j}$ is the \mFpa\ at
  $\nu_{j}$, and $\sigma_{j}$ is the estimated standard deviation of
  the data $D_{j}$ (see Appendix \ref{sec:app:ind} for more details of
  the estimation of $\sigma_{j}$).  The value of $\chi^{2}_{r}$ is
  calculated for both models.  Note that the \mFpa\ are only
  approximately normally distributed, and so caution must be used in
  interpreting $\chi^{2}_{r}$. Since $N>>k$, $\chi^{2}_{r}$ is
  understood as a measure of the average scaled variance of the model
  fit to the data. Within the approximations of the analysis, the
  $\chi^{2}_{r}$ measures show that reasonable fits are obtained for
  Model $M_{2}$ in all cases, excepting the loop footpoints.  The
  relatively low value of $\chi^{2}_{r}$ for these locations indicate
  that probably the size of $\sigma_{i}$ are over-estimated.}

Table \ref{tab:parameters} gives the parameter estimates and 68\%
credible intervals of the parameter values for model $M_{2}$, {\BF the
  maximum value of the ratio $G(\nu)/P_{1}(\nu)$ and its location}.
The \PL\ indices for all regions and both wavebands lie in the
range 1.8 to 2.3, although there is no consistency between one
waveband and another.  For example, the 171\AA\ and 193\AA\ loop
footpoint results both have the same \PL\ index, but the other
regions do not.  However, it should be noted that there is no {\it a
  priori} reason why the \PL\ indices for a region observed in two
different wavebands should be the same.  This is because each waveband
is designed to examine different features on the Sun in broadly
different temperature ranges, and may be imaging {\BF different parts
  of the same temperature-dependent physical process.  Also,
  line-of-sight effects may be different in different wavebands (and
  locations).}  Hence, the emission observed in one waveband need not have
the same statistical properties as any other.

The ratio of the two components $G(\nu)/P_{1}(\nu)$ measures the
relative contribution of the ``excess'' emission over the
\PL\ component.  The frequency $\nu_{max}$ is the frequency at which
this ratio is a maximum, and indicates the most likely frequency at
which the signal of the $G(\nu)$ contribution is strongest.  The
ratios maximize in the range $1.90 \le \nu_{max} \le 4.72$ mHz.  This
is in the frequency range suggested by the moss observations of
\cite{2003ApJ...595L..63D}.  If Model 2 is indeed a reasonable model
of the \Fpa\ in the regions studied, then the $G(\nu)$ contribution is
consistent with previous observations.

Note that neither models $M_{1}$ nor $M_{2}$ feature any narrow-band
oscillatory signals that may be present in the regions studied.  Such
features were not modeled in this study as their contribution to the
overall signal is relatively small and localized to specific
frequencies when compared to the general shape of the power spectra
over the approximately three orders of magnitude (in frequency)
studied here.  The question of how best to determine the presence or
otherwise of narrow-band oscillatory power in \Fps\ is of importance
to the study of coronal seismology, and is addressed in Sections
\ref{ssec:corseis} and \ref{sec:oscdetect}.

\begin{figure}
\label{fig:spatialdistribution:preference}

\caption{Spatial distribution of model preference}
\end{figure}

\begin{figure}
\label{fig:spatialdistribution:powerlawindex}

\caption{Spatial distribution of the power law indices.  Power law
  indices are taken from either model $M_{1}$ or $M_{2}$.  Models are}
\end{figure}

\begin{figure}
\label{fig:spatialdistribution:gaussianposition}

\caption{Spatial distribution of the Gaussian position}
\end{figure}

\begin{figure}
\label{fig:spatialdistribution:gaussianwidth}

\caption{Spatial distribution of the Gaussian width}
\end{figure}

\begin{figure}
\label{fig:spatialdistribution:gaussianamplitude}

\caption{Spatial distribution of the Gaussian amplitude}
\end{figure}

\begin{figure}
\label{fig:powerlawindicescc}

\caption{Two dimensional histograms of the power law indices in every
  pair of AIA EUV channels.  The red line indicates the line where the
  power law indices are equal. Table \ref{tab:powerlawindicescc} shows
the corresponding cross correlation coefficients and the likelihood
that such a cross-correlation coefficient occurred randomly.}
\end{figure}

\section{Discussion}\label{sec:discuss}
The presence of a \PL\ \PS\ has implications for the detection of
narrow frequency band oscillations against such a background emission,
for both non-automated and automated methods.  These are discussed in
Sections \ref{ssec:corseis} and \ref{sec:oscdetect}.  Further, the
\PL\ Fourier spectrum in coronal emission poses questions about how
that emission is formed.  Section \ref{ssec:excess} discusses possible
reasons for the localized contribution component $G(\nu)$.  A
hypothesis regarding the formation of this \PL\ spectrum is
presented in Section \ref{ssec:nplps}.

\subsection{Nature of the \protect\PL\ \protect\PS}\label{ssec:nplps}
{\BF Power-law \PS\ have been
  observed in other solar phenomena.  As was mentioned above,
  \cite{2014AA...563A...8A} observe \PLs\ in the integrated
  emission of small portions of active regions and the quiet Sun as
  observed in the 195\AA\ passband images from EIT over the frequency
  range 0.01 - 1 mHz. Also, \citet{gupta2014} showed \PL\ \PA\ in the
  intensity at six single points in AIA 171\AA\ coronal plumes extending
  over the frequency range $0.3\rightarrow 4.0$ mHz.  Lower in the
  solar atmosphere, \cite{2008ApJ...683L.207R} show the presence of
  \PL\ Fourier \PA, in the range 7-20 mHz, in the Doppler velocity of
  the chromospheric Ca II 854.2 nm line as observed by the
  Interferometric Bidimensional Spectrometer (IBIS).  Further out in
  the solar atmosphere at 2.1 $R_{sun}$, \cite{2008ApJ...677L.137B}
  show the presence of \PL\ \PA\ in Ultraviolet Coronagraph
  Spectrometer observations of the intensity of Lyman-$\alpha$ in the
  frequency range $2.6\times10^{-6}\rightarrow1.3\times10^{-4}$ Hz.}

{\BF The coronal regions studied here also exhibit \PL-like \PA\ in the
approximate frequency range $0.05\rightarrow 10$ mHz.  Under the
modeling assumptions of models $M_{1}$ and $M_{2}$, the first
contribution to the observed power spectrum is a \PL\ (which is
detectable in all regions up to about 2 mHz)}.  It is present in both
wavebands studied and in all the regions, and so parsimony suggests
that the mechanism of its creation is both similar and ubiquitous in
all {\BF similar} parts of the solar atmosphere {\BF , that is quiet
  Sun, moss, loop footpoints and above sunspots}.  Further,
\cite{2014AA...563A...8A} show that a \PL\ \PS\ is present in
time-series of coronal active region and quiet Sun emission over the
frequency range 0.01 - 1 mHz, overlapping with the frequency studied
in this paper.  We hypothesize that the \PL\ \PS\ in the coronal
171\AA\ and 193\AA\ emission is due to the sum of a distribution of a
large number of events having different amounts of emission.
\cite{2011soca.book.....A} describes how a \PL\ \PS\ may obtained from
such a distribution, and that argument is outlined below.

The derivation of the \PL\ \PS\ Equation \ref{eqn:finalfps} shares
many similar features with the nanoflare model of {\it energy}
deposition in the solar atmosphere. In the nanoflare model advanced by
\cite{1988ApJ...330..474P}, large numbers of small magnetic
reconnection events convert the energy in the magnetic field into
energy that heats the corona.  Nanoflares are hypothesized to be small
($10^{24}-10^{27}$ ergs) yet ubiquitous in the corona.
\cite{2013ApJ...771...21W} and \cite{2013ApJ...770L...1T} analyzing
narrowband 193\AA\ images taken by the Hi-C instrument on board a
sounding rocket demonstrate the heating of small scale coronal
structures that appear to be consistent with scenarios of nanoflare
heating due to reconnecting magnetic loops.  However, the nanoflare
occurence rate, or their energy distribution for the entire solar
atmosphere is as yet unknown.

Any modeling or theoretical effort to infer the occurence rate and
energy distribution of energy deposition events from the observed
emission \PL\ \PS\ depends strongly on three factors that influence
the observation.  Firstly, many different loops lie along the
line-of-sight in the corona.  Each of these may have different
emission measures, temperatures and densities.  The location of energy
deposition may not be the same for each loop
\citep{2003A&ARv..12....1W, 2006SoPh..234...41K}.  This implies that
attempts to work back from the observed emission to the underlying
energy deposition must properly account for these line-of-sight
effects.

Secondly, the amount of energy and the wavelengths at which it is
radiated is a highly nonlinear function of the energy deposited in
each event, depending on the existing plasma properties and the
location of the energy deposition in the loop structure.  This must
all be consistently modeled on each loop and summed over multiple
loops to mimic the line-of-sight effects, as noted above.  For
example, \cite{2011ApJ...743...23M} use a full three-dimensional MHD
model to simulate an evolving solar atmosphere and its emission
\citep{2011AA...531A.154G}.  \cite{2008ApJ...682.1351K,
  0004-637X-752-2-161} have developed a zero-dimensional model of
individual strands in a coronal loop that describes the strand's
average temperature, pressure, and density.  Such models can be run
quickly and so can be used to model hundreds of coronal loops.

Finally, the AIA instrument filter response at the time of the
observation must also be understood and modeled {\BF in} order to
understand which emission lines contribute to the final observed
image, and ultimately the \Fps\ for each region.
\cite{2010AA...521A..21O} assume differential emission measure curves
for different regions on the Sun (active region, quiet Sun, coronal
hole, flaring flasma) and calculate synthetic spectra.  These
synthetic spectra were convolved with the effective area of each
channel, in order to determine the dominant contribution in different
regions of the solar atmosphere.  However, as noted by
\cite{2011ApJ...743...23M}, the differential emission measure curves
selected by \cite{2010AA...521A..21O} are not guaranteed to
accurately describe the conditions actually present, and so the
fractional contributions of each line in the AIA waveband are
estimates only.

These three factors will all have an effect on the inferred properties
of the energy deposition rate.  The observed \PL\ \PS\ here, coupled
with the emission event hypothesis and a modeling effort that takes
account of the factors above, offers a new way to constrain the energy
deposition rate in the solar corona.  Under the hypothesis, the energy
deposition rate can be explored through the distribution of the number
of events along the line-of-sight and their temporal dependence,
expressed through the parameters $\alpha_{E}$ and $\gamma$ in Equation
\ref{eqn:energydistrib}.

\section{Conclusions}\label{sec:conc}
Time series of AIA EUV images show approximately power
law like properties at longer frequencies, tailing off to a constant
value at higher frequencies.  A small fraction of time series also
exhibit a narrow band Gaussian contribution over and above the
background power law plus constant \Fps\

.  {\BF It has been shown that the detection of narrow band
  oscillatory power in the practice of coronal seismology must take in
  to account the presence of the \PL\ like \PS.  Detection thresholds
  should also increase in order to reduce the number of false positive
  detections in the presence of \PL\ like \PS.} The power spectra have
been modelled using three components: a \PL\ \PS, a lognormal
component, and a constant background.  It is hypothesized that the
\PL\ component is due to the sum along the line of sight of a
large number of statistically similar energy deposition events.  It is
also hypothesized that the lognormal component is due a combination of
the channeling of oscillatory power from lower to higher levels of the
solar atmosphere and lower temperature emission lines in the AIA
171\AA\ and 193\AA\ bandpasses.

%% If you wish to include an acknowledgments section in your paper,
%% separate it off from the body of the text using the \acknowledgments
%% command.

%% Included in this acknowledgments section are examples of the
%% AASTeX hypertext markup commands. Use \url without the optional [HREF]
%% argument when you want to print the url directly in the text. Otherwise,
%% use either \url or \anchor, with the HREF as the first argument and the
%% text to be printed in the second.

\acknowledgments

We are grateful to the developers of SSWIDL
\citep{1998SoPh..182..497F}, IPython \citep{ipython}, SunPy
\citep{mumford-proc-scipy-2013}, PyMC
\citep{Patil:Huard:Fonnesbeck:2010:JSSOBK:v35i04}, matplotlib
\citep{Hunter:2007} and the Scientific Python stack for providing data
preparation, manipulation, analysis and display packages.  This work
was supported by NASA award NNX13AE03G S01 funded through NASA ROSES
NNH12ZDA001N-SHP, and by a NSF Career grant 1255024 (JMA).

%% To help institutions obtain information on the effectiveness of their
%% telescopes, the AAS Journals has created a group of keywords for telescope
%% facilities. A common set of keywords will make these types of searches
%% significantly easier and more accurate. In addition, they will also be
%% useful in linking papers together which utilize the same telescopes
%% within the framework of the National Virtual Observatory.
%% See the AASTeX Web site at http://aastex.aas.org/
%% for information on obtaining the facility keywords.

%% After the acknowledgments section, use the following syntax and the
%% \facility{} macro to list the keywords of facilities used in the research
%% for the paper.  Each keyword will be checked against the master list during
%% copy editing.  Individual instruments or configurations can be provided 
%% in parentheses, after the keyword, but they will not be verified.

{\it Facilities:} \facility{SDO (AIA)}.

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc.

\appendix\section{Data preparation and model fitting}\label{sec:app:data}
The data was prepared for analysis using the SolarSoft / IDL routines
READ\_SDO and AIA\_PREP.  These procedures convert the downloaded
level 1.0 FITS cutout files to level 1.5 FITS files.  This involves
translating, scaling and rotating the images so they have the same
sun-center, image scale with solar north in the same direction.  All
further processing from this point was performed in the SunPy 0.6
(http://www.sunpy.org) data analysis environment.  Images were
de-rotated to compensate for solar rotation, based on the center of
the field of view of each image.  After de-rotation, each image layer
was coaligned with a reference layer, defined to be half way through
the dataset, in this case at approximately 03:00 UT.  This process is
required since the solar de-rotation function used does not fully
compensate for all the solar rotation. Image coalignment was
implemented using a template matching technique \citep{lewis1995fast}
implemented by the {\it scikit-image} \citep{Vanderwalt2014} routine
{\it match\_template}.  A rectangular template sub-image with sides
one half the size (in pixels) of the reference layer is defined.  The
coalignment algorithm finds the location of the best match of this
template with each layer.  The image layer is then shifted (using
sub-pixel shifts) according to the location of the best match of the
template with the image.  This process removes residual bulk motions
of the images.

In all AIA channels except 131\AA\ the measured intensity of the data
was used to calculate the shifts required for co-alignment.  For
131\AA\ the square root of the measured intensity was used to
calculate the shifts.  This was necessary due to the appearance of
small, localized intense brightening at time ??? - ??? that caused the
template matching technique to fail in 131\AA\ when using the measured
intensity.

The image shifts caused by the co-alignment routine are shown in
Figure \ref{fig:coalignment}.  Note that there is rapid and sudden
change in the shifts around ???-???, corresponding to ??? frames in
each channel's data cube.  Examination by eye of the movies of the
data cubes before this co-alignment step shows that there is a
noticeable shift in all features in the field of view as suggested by
the image shift plot Figure \ref{fig:coalignment}.  Examination by eye
of the movies of the data cubes after the co-alignment shifts have
been applied shows no noticeable shift, leading to the conclusion that
the template matching routine has corrected for the shift adequately.
%% As a final check, the analysis detailed in Section \ref{???}  was
%% applied to the co-aligned datacubes with the final ??? frames removed.
%% Excluding these frames did not significantly affect the final results
%% or conclusions.

\subsection{Fitting the data}\label{sec:app:fit}
Section \ref{???} describes two models of the Fourier power spectrum.
These models are fit to the observed Fourier power spectrum as
described in Section \ref{}.  \Fpa\ models are fit by first generating
an estimate of the model parameters.  This estimate is used as an
initial guess to a fitting routine.

The initial guess for model \ref{eqn:pwrlaw} is generated as follows.
A power law index $n$ is estimated using the observed Fourier power
spectra at the fifty lowest non-zero frequencies using the estimation
approach detailed in Section \ref{sec:app:est}.  For the six hours of
data studied, this covers the frequency range $0.046\rightarrow2.36
mHz$.  The fifty lowest non-zero frequencies are used as this is
typically the frequency range where the observed \Fps\ is dominated by
a power-law like behavior.  An estimate for the power law amplitude
$A_{p}$ is found be averaging the observed power at the first five
non-zero frequencies, covering the frequency range
$0.046\rightarrow0.28 mHz$.  An estimate for the background constant
value $C$ is defined by taking the exponential of the average of the
logarithm of the observed power over the last fifty Fourier
frequencies, covering the frequency range $39.35\rightarrow41.6 mHz$.
Averaging over the logarithm of the power reduces the chance that the
estimate of $C$ will be strongly biased by the largest power in the
frequency range.

The initial guess for model \ref{eqn:pwrlawbump} is generated as
follows.  Estimates for $n$, $A_{p}$ and $C$ are generated as
described in the previous paragraph.  An estimated power spectrum
$P_{est}$ is then calculated using these estimates and model
\ref{eqn:pwrlaw}.  If there at least ten positive differences between
the observed Fourier power and estimated power spectrum $P_{est}$ in
the frequency range $0.97 \rightarrow 9.31 mHz$, then these positive
differences are used to estimate the parameters of the Gaussian,
equation \ref{eqn:bump}.  The parameters $A_{G}$ and $\beta$ are
estimated as the maximum value of the difference and the logarithm of
the Fourier frequency at which this maximum value occurs,
respectively.  The width $\sigma$ is estimated by calculating the
standard deviation of the positive differences with respect to the
logarithm of the frequency.  If there are less than ten positive
differences in the required frequency range, it is assumed that there
is a very small Gaussian contribution to the observed \Fps.  In this
case, $\ln A_{G}=-100$, $\beta$ is fixed to be half way between the
logarithm lower and upper estimated Gaussian frequency range $0.97
\rightarrow 9.31 mHz$, and $\sigma = 0.1$.

Model fitting was initialized using the initial estimates described
above.  Models were fit by varying the value of model parameters in
order to find the maximum of the log of likelihood of the data given
the model.  If the time-series data are subject to a Gaussian noise,
then the Fourier power components are exponentially distributed.  Let
us assume that $I_{j}$ is the observed Fourier power at the frequency
indexed by $j$, and $S_{j}$ is the model Fourier power at that same
frequency.  The probability of an observed power $I_{j}$ given a model
power $S_{j}$ is given by
\begin{equation}
p_{j}\left(I_{j}|S_{j}\right) =
\frac{1}{S_{j}}\exp\left[-\frac{I_{j}}{S_{j}}\right]
\label{eqn:freqprob}
\end{equation}
If there are $N$ non-zero Fourier frequencies then the likelihood over
all frequencies is
\begin{equation}
L=\prod_{j=1}^{N}p\left(I_{j}|S_{j}\right)
\label{eqn:freqprob:all}
\end{equation}
The Nelder-Mead method was used to vary model parameters in search of
the maximum of the log likelihood.  The full code used (and the
history of its development) is availble as a git repository at
https://github.com/wafels/rednoise/tree/perpixel.

\subsection{Estimating the power-law index in a power law power
  spectrum}\sec{sec:app:est}
Equation \ref{eqn:freqprob:all} describes the likelihood of the
observed Fourier power given a model for that power.  We assume that
the observed power law spectrum can be modeled by a simple power law
over all frequencies
\begin{equation}
S_{j} = Af_{j}^{-n}
\end{equation}
where $A$ is a constant, $f_{j}$ is the $j$'th Fourier frequency, and
$n$ is the power law index we wish to estimate.  An estimate can be
derived using Bayes' theorem. Using the likelihood function derived
above (\ref{eqn:freqprob:all}), and assuming the prior $p(A)\propto
A^{m}$ for some value of $m$, then 
\begin{equation}
p(n) \propto \int_{0}^{\infty} p(A) L dA.
\label{eqn:bayes}
\end{equation}
On integration,
\begin{equation}
\ln p(n) \propto n\sum_{j=1}^{N}\ln f_{j} - (N-m-1)\ln\left[\sum_{j=1}^{N}I_{j}f_{j}
\label{eqn:pnlog}
\end{equation}
For the purposes of this investigation, $m=0$ and an estimate to $n$
was found by finding calculating the value of $\ln p(n)$ in the range
$n=0\rightarrow4$ with steps of 0.01 and finding the value of $n$ that
maximized $\ln p(n)$.

\subsection{Calculating the 95\% confidence level using the $\chi^{2}$
analog}\label{sec:app:chi}

\subsection{Acceptable parameter limits}\label{sec:app:limit}

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% AASTeX provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\bibliography{references}
\end{document}